The Eclipse Integrated Computational Environment

### 2019-07-28 src/conclusions.tex
Moved conclusions to separate file.
Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

\section{Motivation and Significance}\label{motivation-and-significance}

In previous work, Billings et al. interviewed modeling and simulation subject
matter experts to compile a list of requirements for implementing and using
these kinds of applications. In the process, they discovered that many of the
difficulties inherent in using high-performance modeling and simulation
software fall into five distinct categories \cite{billings_designing_2009}.
These activities, detailed in Section \ref{workflow-model}, include (1)
creating input, (2) executing jobs, (3) analyzing results, (4) managing data,
and (5) modifying code. There are many tools that address these problems
individually, but the same research found that the excess number and
specialization of these tools also contribute to the learning curve.

Previous efforts to address these five issues have resulted in general-purpose
scientific workflow tools like Kepler \cite{ludascher_scientific_2006} or
myopic tools that satisfy only a single set of requirements for a single piece
of software or a single platform. These are opposite extremes, but a
middle-of-the-road solution is also possible. A workflow engine could be
developed that limits its scope to high-performance computing (HPC) and to the
set of possible workflows associated with the five previously mentioned
activities. With only minor additional development, a rich application
programming interface (API) could be exposed so that highly customized
solutions could still be made based on this limited workflow engine.

It is not clear which, if any, of these solutions is better than the
others, and practical requirements will ultimately dictate the path of a
project's progress. This work considers a middle ground solution and
presents the Eclipse Integrated Computational Environment (ICE) as proof
that it is possible to create such a system. Specifically, the work
described here shows that

\begin{itemize}
\item
  modeling and simulation activities can be described in a succinct
  workflow model (see \S \ref{workflow-model}),
\item
  an architecture for such a workflow system can satisfy the model of
  workflows in an extensible way (see \S \ref{software-architecture}), and 
\item
  such a system is applicable to a suite of problems in energy science,
  including virtual battery simulations and additive manufacturing, 
  among others (see \S \ref{illustrative-examples}).
\end{itemize}

This section concludes with an introduction to the ICE workflow model. Section \ref{software-description} details the software
from an architectural perspective, and Section \ref{illustrative-examples}
provides a set of comprehensive examples. A
presentation of the impact is included in Section \ref{impact}, and details on obtaining sample 
code are provided in Section \ref{sample-code-tutorials-and-other-resources}.

\subsection{Workflow Model}\label{workflow-model}

ICE's workflow model is based on making it easier for scientists to create
input, launch jobs, analyze results, manage data, and modify code. Many
scientists would most likely find these activities difficult for all codes with
which they lack experience, whereas with their own codes---or those with which
they are most familiar---these tasks may be so simple that they are taken for
granted. Any particular combination of these activities across one or more
scientific software package or code results in a unique workflow. Such a
workflow is normally, but not always, requested by a human user and
orchestrated by a workflow management system.

The most obvious workflow for any individual simulation code or
collection of codes is to string the activities together, where the user's
workflow is to create the input, launch the job, perform some analysis,
and manage the data---possibly modifying the code in the process. However, 
there are many other combinations, including re-running jobs with
conditions or modifications or analyzing someone else's data.\footnote{The
authors have identified many unique combinations that
define workflow ``classes.'' When possible, every effort is made to give the
classes colloquial names such as ``The Re-Run'' or ``The Graduate Student.''}

\textbf{Creating input} is the process of describing the physical model
or state of a system that will be simulated. This could include creating
an input file(s) or making calls to an external process to configure a
running program. In most situations, a computational scientist will
modify existing input or create new input from a template. ``Input''
generally includes run time parameters for the simulation framework
(e.g., tolerances); configuration options (e.g., data locations, output
locations, module configurations); properties of the materials to be
simulated; and a discretization of the simulation space (e.g., mesh,
grid, particle distribution). The collection of all required input can
be quite large and can go by many names, including ``input set,''
``input package,'' ``problem,'' or, simply, ``input.'' Often, the set of
input files will be described in a ``main'' input file that acts as a
kind of manifest to describe---and provide links to---all necessary
information for a given problem.

In this work, it should be assumed---unless otherwise noted---that
``input'' refers to the entire set of input, not to a single file.

\textbf{Executing jobs}, or ``running the workflow'' in this context, is
the process of performing calculations using a simulation code or
framework based on known variables from the input. These are typically run locally for small jobs or for development. Large
simulations, on the other hand, typically require a large amount of
hardware resources. These resources are usually off-site (i.e.,
physically unavailable to the user) and are accessed remotely
through Secure Shell (SSH) connections or similar protocols. Remote
execution requires moving the input in advance of the execution and
copying or moving the output to the user's machine. In many cases,
though, the output is too large to move to the user's local machine.

Local and remote jobs are often monitored to ascertain a job's status.
This monitoring could be a simple check as to whether or not the execution
has completed, or it could involve monitoring the output of individual
quantities to examine the calculation state. The latter is often used to
detect calculation errors that will result in incorrect results. If such
problems are found, the job is typically cancelled (``killed'') to save
compute resources and is then re-run later.

Local jobs in ICE are executed using standard Java system calls. Remote jobs
are launched only through SSH connections on remote machines. This includes
direct SSH command execution on clusters and proxy connections through a pilot
service on large leadership-class supercomputers. Services such as Globus GRAM
and Bosco are not supported, but the SSH command execution includes extensive
support for numerous batch and queuing systems. The Eclipse Parallel Tools
Platform (PTP) is used to create all remote SSH connections, regardless of the
target machine \cite{tibbitts_integrated_2009}.

In this work, it should be assumed---unless otherwise noted---that
``executing a job'' includes monitoring that job in one or more ways,
possibly including real-time updates to visualizations. It is also important to
note that executing a job is not the same as executing a workflow. Executing
jobs specifically refers to launching simulations, whereas executing a workflow
could be something different such as generating input or postprocessing
results. This is an important distinction because ICE's workflows are all
executed locally, but simulations and work can be distributed remotely
depending on the implementation of the workflow plugin.

\textbf{Analyzing results} includes executing special jobs to transform
data in one or more prescribed ways and producing artifacts with
scientific significance from the transformed data. This could include, for
example, postprocessing results and visualizing the new data with
dedicated visualization tools. For many types of scientific computing,
this includes viewing the results of a simulation on a mesh or grid and
extracting publication-quality images or movies from that data. Other
cases might include analyzing results in preparation for follow-on
simulations or performing feature extraction, classification, or
activities for machine learning and data mining.

Although this is similar to executing a job, it is distinctly
different because the activity changes focus to satisfy the needs of a
human operator. Simple data reduction, where the exact reduction is
known, certainly qualifies as executing a job; however, analysis of
modeling and simulation results is far from simple data reduction and is
generally far more interactive for scientists.

\textbf{Managing data} includes moving, copying, storing, sharing, or
otherwise interacting with data for or from simulations. This activity
is the most pervasive because each of the other activities requires
interacting with data in some way. In many cases, though, data is still
managed for its own purposes, without performing a simulation,
generating new input, or analyzing results. Examples include archiving
data, packaging data for publications, and updating values manually or through
scripts (often in light of new information from publications).

\textbf{Modifying code} is not typically considered a part of a
scientific computing workflow. However, modeling and simulation use
cases often require users to explicitly modify code before execution or to
issue special build instructions. This is true, for example, with the
computational fluid dynamics code Nek5000, which requires modification of code
before compilation and special build instructions using the \textit{makenek}
shell script \cite{the_nek5000_team_nek5000_2014}.
Many scientists consider ``their workflow'' to be re-running
software after modifications for purely exploratory purposes. This might
be required if the model the author is modifying cannot be
configured directly as part of the input but can be easily accomplished by
manipulating the source code.

\subsection{Related Work}\label{comparison-to-other-models}

ICE's model of workflows differs significantly from many other
efforts in workflow science because it defines workflows in terms of high-level
activities that are meant to guide workflow implementers and users. Many other
workflow models in the literature define a workflow as a collection of
computing processes. For example, Yu and Buyya define grid workflows as ``a
collection of tasks that are processed on distributed resources in a
well-defined order to accomplish a specific goal'' \cite{yu_taxonomy_2005}.
Others, such as Pizzi et al., subscribe to similar definitions
\cite{pizzi_aiida:_2016}. This ``process'' view is acceptable where the
workflow is static and does not require additional human input or ``human in
the loop'' behavior after all the initial human input is provided. However,
workflows within ICE are fully interactive with regular callbacks to humans. It
is simpler to discuss ``activities'' than it is to create a distinction between
``human processes'' and ``computer processes.'' Focusing on activities over
processes (human or computer) also has the benefit of removing concrete
elements such as hardware or software properties that distract from details of
workflows and workflow management systems. That is, considerations such as
memory usage and raw performance are important, but questions about the
abstract workflow or what the workflow management system should do are far more
important in this context.

There are other workflow engines that have their own concepts of activities,
with various similarities and differences compared with ICE's model. For
example, Taverna has a concept of activities \cite{wolstencroft_taverna_2013}.
Activities in Taverna are defined as invocable activities within a workflow,
which is closely related to a lower-level class in ICE called Actions (see \S
\ref{item}). On the other hand, ICE's activities loosely describe what can be
done conceptually and the workflows that are implemented provide the Actions
that carry out these activities.

ICE and Taverna can both natively execute Java code, but ICE 2.0 cannot
natively invoke Web Service Definition Language (WSDL) web services. ICE uses
regular Java-based libraries to interact with web services, such as Jersey, and
requires that developers write the code to do this directly. Shim services in
Taverna are provided in ICE as separate services in the framework but are
otherwise very similar.

Chiron, an algebraic workflow engine, has a concept of activities that is
closer to ICE's than Taverna's but still very strongly related to
implementation \cite{ogasawara_chiron_2014}. Activities in Chiron are
combinations of the program or expression to execute, plus the input schema,
plus the output schema. This is a close match to the design of the Item class
in ICE. ICE's Items (see Section \ref{item}) implement the high-level
activities, such as launching a job or creating a model. The major difference
between Chiron's activities and ICE's Items appears to be scope. A Chiron
activity such as Map or Reduce would normally be implemented as an Action in
ICE, whereas an Item in ICE can be tasked with executing multiple Actions to
enact a workflow.

One significant difference between ICE and many other systems is that it uses a
language-based approach to defining workflows but with a standard language,
Java, as opposed to a custom workflow definition language. Pizzi et al. do this
with Aiida and Python, and the Fireworks Workflow Engine also uses Python. The
advantage of this approach is that compilation is not a problem. ICE also
differs from other workflow engines because it is specifically focused on
modeling and simulation instead of grid workflows. Grid workflows are almost
always defined as workflows based on trees without humans in the loop, but
modeling and simulation workflows are much more interactive and undirected
\cite{billings_toward_2017}.

\section{Software Description}\label{software-description}

ICE was specifically created to address hands-on workflows for scientists, as
described in \S \ref{illustrative-examples}. Users download and execute ICE
locally, and ICE in turn orchestrates local or remote workflows as required. It
provides a comprehensive workbench for modeling and simulation that includes
tools for workflows, visualization, data management, and software development.

\subsection{Software Architecture}\label{software-architecture}

\begin{figure}[htbp] \centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ICE-arch-1.jpg}
\caption{High-level architecture of ICE showing the relationships and rough
division of responsibility between the five primary components of ICE. The ICE
Core is responsible for the workflow management and delegations instructions
and data to the Item and data structures, respectively. User input is received
from the client or through headless interactions, while persistence of large
amounts of data is delegated to a persistence service. The ``1'' and ``*''
represent the cardinality of the aggregation relationship.}
\label{highlevel-arch}
\end{figure}

Workflows and tasks in ICE are not explicitly treated as trees or
directed acyclic graphs (DAGS), as is common with grid workflow tools
\cite{yu_taxonomy_2005}. Instead, ICE's design is inspired by representational
state transfer (REST), and the workflow engine is implemented as services in an
Eclipse Rich Client Platform (RCP) application
\cite{fielding_architectural_2000} \cite{mcaffer_eclipse_2010}.

Figure \ref{highlevel-arch} shows the five primary components of ICE and their
relationships to each other. Users initiate requests to create, edit, update, or
delete workflows from the ``ICE Client'' (the workbench) or through headless
interactions using web or language programming interfaces. The list of available
workflows that can be created is provided dynamically to the ICE Client by the
``ICE Core,'' which acts as a server and is the primary component responsible
for workflow orchestration. Workflow information is provided dynamically because
it often changes at run time based on both the configuration of available
workflow components in the registry and on persisted workflows users have
saved in their workspace. Workflow task enactment and execution is performed by
services called ``Items.'' Information about workflows is provided to the Client
by the Core through common data structures rooted in a stateless ``Form'' tree.
Forms describe the workflow and provide all the necessary information to
understand \emph{what} should process the workflow (but not \emph{how} it should
be processed). Once users modify the description of the workflow in the Form to
provide their specific details, the Client dispatches a request to the Core to
modify and/or process the workflow. The Core then uses the information from the
Form to perform a service lookup to identify the correct Item to process the
workflow.

\subsubsection{ICE Core}

\begin{figure}[htbp]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ICE-arch-2.jpg}
\caption{Architecture of the ICE Core and its dependencies. The ICE Core class
handles communications between workflow Items and common data structures but
delegates construction of workflow Items to builders.}
\label{core-arch}
\end{figure}

The primary purpose of the ICE Core is to orchestrate workflows and services for
the platform. Figure \ref{core-arch} shows the primary classes that interact
with the ICE Core. It is the ``core'' control component of ICE where all other
component interactions converge. It is the middle tier of ICE's architecture and
brokers communications between Items and Clients (graphical or not) by sharing
Forms that contain workflow data. The ICE Core class delegates construction and
management of Items to the Item Manager.

The Item Manager acts as a container and factory for Items created by the ICE
Core. Items are constructed using the Builder pattern, with concrete
construction handled by realizations of the ItemBuilder interface.

ICE 2.0 includes a graphical client based on the Eclipse Workbench, which is
common for Eclipse applications. The workflow management capabilities of the
core can be accessed through languages and a web-based programming interface as
well (see Section \ref{framework}).

\subsubsection{Items}
\label{item}
\label{item-states}

\begin{figure}[htbp]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ICE-arch-3.jpg}
\caption{The class hierarchy of ICE Items showing the relationships between the
Item, its subclasses, and services that it uses. Vibe Launcher and MOOSE
Launcher are examples of special Job Launch Items for executing jobs, and their
applications are discussed in \S \ref{illustrative-examples}.}
\label{item-arch}
\end{figure}

Each workflow Item in ICE is an independent service, and each workflow type is a
subclass of the Item base class or another Item subclass. Items are registered
dynamically through a service registry in ICE (see Section \ref{framework}) and
provide collections of workflow data ---Forms--- for the ICE Core and Client.
Since ICE's design is highly object oriented, it is easiest to think of the Item class
as a description of an abstract workflow and an Item object (an instance of the
class) as a concrete workflow with all required execution details specified in
its Form.

Figure \ref{item-arch} shows the class hierarchy of the Item class and its
collaborators.
Individual components of workflows (i.e., workflow ``tasks'' or ``nodes'') are
either encoded directly in the workflow's subclass of Item or provided as
``Actions'' that are dynamically registered with an ``Action Factory'' and
obtained at run time. Common data structures that are shared across Items, the
Client, and the Core are provided by the data structures component. Table 1
describes the differences between Items, Actions, and Forms.\footnote{An
upcoming update to the API will include the formal introduction of IWorkflow,
IWorkflowTask, and IWorkflowEngine interfaces to bring ICE's API language closer
to other systems such as Triquetrum. However, this must be done carefully to
preserve backwards compatibility.}

\begin{table*}[t]
\begin{tabularx}{\textwidth}{|l|X|l|}
\hline
Class & Class Description & Object Description\tabularnewline\hline
Item & Java class with code to execute an abstract workflow. Provides a
Form. & Concrete workflow executor.\tabularnewline\hline
Form & Description and template of the data needed for the Item to
process the workflow. & User-modified workflow data.\tabularnewline\hline
Action & Java class for executing a specific task in the workflow. Used
by Items. & Concrete workflow task executor.\tabularnewline\hline
\end{tabularx}
\caption{Class descriptions for Items, Forms, and Actions.}
\end{table*}

Commonly used subclasses of Item include the Model and Job Launcher classes,
which can be  used to create workflow plugins for input generation and workflow
execution, respectively. Composite Items are special Items that are built of
compositions of other Items. Composite Items are commonly used for workflow
Items that can be executed in parallel or to tightly couple different workflow
tasks for resource and performance management.

All Items in ICE are finite state machines where the states represent
the abstract state of the workflow. For example, when an Item is first
created, it enters the ``Form Ready'' state to indicate that it could,
in theory, be processed after a user reviews it. After that review, it
enters the ``Ready to Process'' state before it is processed and 
the ``Processed'' state after it is processed. There are several additional 
states for errors.

This design is very important. First, it means that all workflows in
ICE, regardless of their actual details and functions, can behave only in
a specific set of known and predictable ways, and this predictability 
simplifies the way the Core manages and interacts with Items. Second, by
formalizing state and error checks, ICE explicitly delineates which workflows 
can be executed from those that must receive additional configuration. 
Finally, it allows developers implementing Items and
Actions to specify by contract what is required before proceeding to the
next task, processing the workflow, or declaring a successful execution.

\subsubsection{Data Structures}

ICE contains a collection of common data structures that are useful for
representing data needed modeling and simulation workflows. Figure
\ref{data-arch} shows some of these classes and their relationship. All data
structures (and Items) in ICE are uniquely identifiable and behave in
predictable ways: they receive and dispatch updates, can be visited to determine
type information, etc. \textit{Components} in the data sense, not the class
sense, represent a special data structure that forms a child of a Form. Forms
themselves are components and hierarchically aggregate other components.

The name ``Component'' in the context of ICE's data structures refers to a
unique and reusable collection of data, not a class or component in the object
oriented sense. Components that hold data do very little of their own work but
strongly scope the interfacial contract of their clients.

\begin{figure}[htbp]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ICE-arch-4.jpg}
\caption{Common data structures and their relationships in ICE.}
\label{data-arch}
\end{figure}

\subsubsection{Persistence and
Workspaces}\label{persistence-and-workspaces}

When workflows are created and modified, ICE saves permanent copies of 
Forms to disk in a special directory called a ``workspace.''
Workspaces can contain projects, folders, and files, including data,
code, input, and output. ICE automatically manages local and remote (or
even local \emph{to} remote) transfers of data files when executing
workflows if the files are detected in the same directory of the
workspace as the workflow itself. For example, when executing a remote
job, ICE will automatically move the input file if it is specified in
the workflow and available in the project directory of the workspace.
Likewise, if the output is small enough (less than 50 megabytes), ICE will
automatically move it back to the local directory. By convention, all paths in
ICE's Forms are relative to the workspace root path. Workspace directories are
specified by the user and are fully functional Eclipse workspaces.

ICE handles persistence using a ``persistence provider.'' The
default persistence provider uses JAX-RS to write Forms to 
XML~\cite{burke_restful_2010}. In principle, other persistence providers 
could replace the XML-based provider since it is registered as a dynamic 
service and a JAXP based provider has been used in the past.

\subsubsection{Relationship to Other Eclipse Technologies}\label{framework}

ICE is an Eclipse RCP application \cite{mcaffer_eclipse_2010} and has a plugin
architecture based on Equinox, which is the reference implementation of the Open
Service Gateway Initiative (OSGi) framework specification
\cite{mcaffer_osgi_2010}. It is most appropriate to think of Eclipse ICE as an
alternative flavor of any other Eclipse development environment because like
those environments it is just a collection of Eclipse plugins with a
well-defined purpose and brand. ICE uses more than 1,200 additional packages
from the Eclipse ecosystem to provide services like language support and
visualization. Each unique element of ICE described in this work---including the
Client, Core, Items, and data structures---is provided as plugins to Equinox.
Most plugins are managed dynamically and provided as services that can be
obtained as dynamic OSGi Declarative Services. File input/output in ICE, with
only a few exceptions, interacts with the RCP's Resources plugin and the
standard Eclipse Workspace. This also includes remote resources that are
managed with Eclipse PTP.

ICE plugins are created using normal Eclipse development tools and are installed
into either the running instance of ICE or published in a newly built version.
In either case, the running platform of ICE can use only the plugins that are
actually installed and at least a small amount of code must be written in Java
for each plugin. This extension method may seem odd to the novice user or
developer, especially if they do not know Java very well, but this design was
chosen because it is an easy, fast, and precise way to create workflows.

ICE provides built-in development tools to automatically generate empty but
ready to compile classes (``stubs'') of plugins that can then be installed into
the framework as services. These tools are based on other code generation tools
in the Eclipse platform. This ``self-hosting'' makes it possible for new users
to create complex, sophisticated workflows very quickly because they are not
required to know how to interact with the framework. Development of new
workflows is streamlined because ICE provides a rich API that exposes all
workflow management functionality, documentation, tutorials, and tools to
further simplify the process. These are shared through the self-hosted
environment by code completion tools, help menus, and documentation overlays.
The proximity of the workflow environment with the development platform
for modeling and simulation codes is also beneficial because, as a self-hosted
entity, workflows can be codeveloped with the code.

ICE can be run as a headless web server with a remote service interface and a web
API. The web API is also used as the primary means of providing real-time 
feedback and monitoring support in ICE, and it is published as an OSGi service
as well. These services can also be consumed as OSGi Remote Services.

There are many situations where configuring workflows graphically or serially is
unacceptable, such as when a very large number of workflows will be
executed or the type of information required is very fine grained. In
these cases, it is often necessary to provide scripts to the workflow
management system. ICE includes the Eclipse Advanced Scripting
Environment (EASE)~\cite{pontesegger_eclipse_2015} for scripting because it
provides a way to script Eclipse RCP projects natively in Javascript,
Jython, and Python. This also makes it possible to extend the
environment by adding Items in these languages. Specifically, scripts in
Javascript and Python are written and executed in simple shell that is part of
the workbench. Calling other languages, such as C or Fortran, is typically
accomplished by adding workflow Items for the executables in these languages
and performing local workflow job launches. C language routines have been
called from ICE through the Java Native Interface (JNI), but this is not a
technique that is widely used with the platform because of its complexity.

\subsection{Functionality}\label{software-functionalities}

The most important function of ICE is to serve as an easily extended
workflow management system for scientists in support of
the activities described previously. In practice, it is most often used as a
combination of workflow management system and development environment
since it contains a significant amount of Eclipse's software development
tooling in addition to the workflow tools. The project has many users,
but it is most heavily used by the development team to support workflow
science and software development for energy science projects. The
development team regularly uses the platform to quickly deploy new
domain-specific workbenches in a matter of hours for small collections
of workflows that are easy to encode.

Outside of the development team, ICE is commonly deployed as a sophisticated
user environment for computational science projects (see Section \ref{impact})
and as a visualization tool. The ICE source code originally contained a
significant amount of visualization support, but at the request of users in the
community, that support was ``spun off'' in early 2016 as the Eclipse Advanced
Visualization Project (EAVP)~\cite{billings_eclipse_2015}.

\section{Illustrative Examples}\label{illustrative-examples}

The role that ICE plays as a workflow tool is best illustrated by the
various ways in which it has been deployed, as shown in the following 
examples.

\subsection{Virtual Battery
Simulations}\label{virtual-battery-simulations}

Pannala et al. developed a Virtual Integrated Battery Environment (VIBE)
as part of their research into safety and performance characteristics of
lithium ion batteries \cite{pannala_multiscale_2015}. VIBE includes ICE
as part of its distribution, and new workflows were added to ICE to enable users
to add multiple types of input, configure the simulation software, and launch
simulations of virtual batteries. Interactive 3D visualizations of the
results were embedded in the launcher so that users can quickly find
their results. Figure \ref{vibe} shows a (simulated) prismatic-cell battery's temperature
distribution during discharge.

\begin{figure}[htbp]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/vibe_20151016.png}
\caption{ICE workbench for VIBE analysis.}
\label{vibe}
\end{figure}

VIBE 1.0 is available as a virtual machine (for convenient deployment) in which the simulation
software and ICE are provided side by side. The VIBE team's more recent 
efforts for VIBE~1.1 include providing the simulation
software in Docker containers so that users can download the latest,
native version of ICE for their machine while simultaneously benefiting
from a smaller virtual machine for the simulator.

\subsection{Multiphysics Simulations with
MOOSE}\label{multiphysics-simulations-with-moose}

The MOOSE Framework is a powerful, easy-to-use multiphysics framework
developed at Idaho National Laboratory \cite{gaston_moose:_2009}. ICE
provides workflow tools for MOOSE as well as specialized class
generation utilities for developing custom MOOSE kernels. Many of the
MOOSE tools in ICE were developed closely with the MOOSE team to
reproduce various aspects of MOOSE's user interface, known as ``Peacock.'' 
Figure \ref{moose} shows an example of the ICE workbench for a simple structural mechanics
problem solved using the MOOSE framework \cite{mccaskey_scientific_2015}.

\begin{figure}[htbp]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ice-moose.png}
\caption{ICE workbench for MOOSE workflows.}
\label{moose}
\end{figure}

There are more than 300 MOOSE-based applications, and it is very easy to create
new ones. The ICE development team uses ICE and MOOSE to quickly solve energy
science problems with HPC resources and to deploy domain-specific workbenches.
ICE provides features for automatic installation, configuration, and
optimization of scientific development environments. In the context of MOOSE,
ICE includes support for automatically downloading and building MOOSE from
MOOSE's GitHub repository. This integration enables users to immediately begin
developing complex multiphysics applications using the preinstalled Eclipse C
Development Tools. Once the new MOOSE-based application is built, it will
automatically work with the MOOSE workflow tools in ICE, although developers can
also create customized workflow tools as needed.

\subsection{Binder Jet Modeling}\label{binder-jet-modeling}

Solid-state sintering of parts printed using binder jetting
significantly increases part strength by decreasing the part porosity
and eliminating voids. However, this process does cause the resulting
product to shrink and warp from its original layout. An ideal near-net-shape 
process would combine binder jetting with solid-state sintering and account
for warpage and deformation in the part's design phase. 
Figure \ref{binder} shows an ICE-based workbench for performing simulations of
this process with visualizations of the pre- and postsimulation
properties of a central body with eight cantilevers. The primary
deformation in this type of geometry is bending or drooping of the
cantilevers due to sintering and thermal creep.

\begin{figure}[H]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/ice-bjm.png}
\caption{ICE workbench for binder jet modeling.}
\label{binder}
\end{figure}

\subsection{Neutron Reflectivity}\label{neutron-reflectivity}

ICE also includes a small utility for simulating neutron reflectivity and
comparing the results with other data \cite{billings_brand_2015}. This 
utility was developed in collaboration with a team at ORNL's Spallation Neutron 
Source to replace an older utility that was originally written in Visual Basic and distributed via Excel macros. The new utility, developed in ICE, is shown in Figure \ref{reflectivity}.

\begin{figure}[H]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/reflectivity-screenshot.png}
\caption{ICE workbench for neutron reflectivity.}
\label{reflectivity}
\end{figure}

\subsection{Quantum Computing}\label{quantum-computing}

As quantum computing grows, the need for sophisticated
software that can use quantum hardware or perform calculations on
simulated quantum hardware becomes more pressing. Humble et al. created a simulator for
adiabatic quantum computers where workflows were added to ICE to support
interactions with the simulator and to process large sets of
quadratic binary optimization problems \cite{humble_integrated_2014}. Figure \ref{jade} shows the workbench for this project.

\begin{figure}[H]
\centering
\baseIncludegraphics{pubs/ice-softwarex-2017/src/images/jaded.png}
\caption{ICE workbench in the Jade Adiabatic Development Environment (JADE) for quantum computing simulations.}
\label{jade}
\end{figure}

%ICE also supports other quantum computing projects, including the
%eXtreme-scale ACCelerator (XACC) effort
%\cite{mccaskey_ornl-qci/xacc_2016}. XACC is a programming framework for
%extreme-scale, post-exascale accelerator architectures that can be
%integrated alongside existing, conventional applications.

\subsection{Nuclear Energy}\label{nuclear-energy}

There are many examples of ICE's role in modeling and simulation
projects for nuclear energy, but for an example of the level of
customization that is possible in ICE, readers are referred to the work outlined in
\cite{billings_domain-specific_2015}. Support for the ``Reactor
Analyzer'' was dropped in ICE 2.1.8, but it demonstrated ICE's capability
to integrate many different nuclear energy tools for complex analyses.

\section{Impact}\label{impact}

The impact of software tools like ICE is difficult to quantify. However, there
are several examples of ICE significantly assisting the development team and
others.

One pressing area of interest and impact is that of interoperability
between workflow systems. Significant prior efforts have been made to
combine large workflow management systems (e.g., Mandal et al.
\cite{mandal_integrating_2007}), but the end goal of gaining the greatest
advantage by using the best capabilities from multiple systems remains unrealized.
ICE's unique perspective on workflows and its well-defined API make it
possible to integrate multiple systems in a straightforward way. This
allows ICE to connect to other workflow environments, such as Triquetrum,
quite easily \cite{brooks_introducing_2016}. Triquetrum, like Kepler in
\cite{mandal_integrating_2007}, is a Ptolemy-based workflow engine
\cite{brooks_triquetrum:_2015}.

It is widely known that tools that enable researchers to be more productive tend
to improve the pursuit of new research. The high extensibility of
ICE and the tools that it combines from the larger Eclipse ecosystem
have made it possible for researchers on the development team to quickly
deploy new simulation environments for their research problems. Other tools created
with ICE might not invent something radically new, but they tend to
streamline interactions within those tools. Many ICE users, and certainly
the development team, have experienced improvements in their software
development efforts because of the tools that ICE provides or have learned
new technologies because access to new tooling was as simple as
installing more plugins through the Eclipse Marketplace.

The ICE development team does not track ICE's user base, as useful as
that would be, because of the extra work involved. However, various
sources such as the VIBE mailing list, ICE's own mailing lists, and
website download statistics suggest that ICE has been used by more than 350
people at one time or another and currently has about 20
``superusers,'' including the development team.

A new cloud-based development tool based on ICE is under development by
RNET Technologies, Inc., out of Dayton, Ohio, in response to a Small Business Innovation 
Research award from the DOE. This web-based version of 
ICE will continue ICE's support for nuclear energy and will integrate with cloud 
computing solutions like Amazon Web Services and ORNL's Compute and Data
Environment for Science (CADES). Additionally, although ICE has not directly led
to any ``spin-off'' companies, ICE source code has been used
in two spin-off projects: (1) EAVP (mentioned earlier) for advanced
visualizations and (2) the Eclipse January project for scientific data
structures \cite{graham_eclipse_2016}. ICE was also one of the founding
projects of the Eclipse Science Working Group.

Eclipse ICE is an open-source project, and the authors welcome and encourage
engagement and contributions from users. Interested parties may inquire by
contacting the corresponding author or visiting the ICE website.

\section{Sample Code, Tutorials, and Other
Resources}\label{sample-code-tutorials-and-other-resources}

The primary resource for information on ICE is the project 
website~\cite{billings_eclipse_2016}. The ``Resources'' menu includes links to
detailed tutorials and user documentation. Examples of how to create
new workflow Items are available at
\url{https://github.com/eclipse/ice/tree/master/org.eclipse.ice.demo}.
Examples of how to use the scripting engine are available at
\url{https://github.com/eclipse/ice/tree/master/examples}. ICE also includes an
extensive suite of unit, integration, and user interface tests, which are
also excellent examples of how to work with the platform. Tutorial and
demonstration videos are available on YouTube at \url{https://goo.gl/nxCzRD}.

\section{Conclusions}\label{conclusions}

Modeling and simulation workflows for scientists differ
greatly from those of experimentalists or those who primarily interact with
grid-based workflow management systems. The Eclipse ICE effort described here 
has been used to address interdisciplinary problems in 
modeling and simulation for energy science. Of particular interest are the
differences in architecture between a workflow management system focused
on modeling and simulation compared with systems focused on grids. ICE's broad
applicability across many topics in energy science suggests
there are opportunities for these systems in general. Finally, one 
interesting avenue of future exploration is coupling or integrating ICE 
with other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus,
which would make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.

>>>>

Modeling and simulation workflows for computational scientists differ greatly
from those of experimentalists or those who primarily interact with grid-based workflow
management systems. The Eclipse ICE effort described here has been used to
address interdisciplinary problems in modeling and simulation for energy
science. Of particular interest are the differences in architecture between a
workflow management system focused on modeling and simulation compared with
systems focused on grids. ICE's broad applicability across many topics in energy
science suggests there are opportunities for these systems in general. Finally,
one interesting avenue of future exploration is coupling or integrating ICE with
other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus, which would
make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.
